---
title: "Ecotox Exploratory Notebook"
output: html_notebook
---

This notebook explores Ecotox data and selects appropriate tests 
and results for the Water 100 Project

# Setup 

## Import libraries 
```{r}
library(tidyverse)
library(dbplyr)


library(readr)
library(here)
library(jsonlite)
library(readxl)
library(webchem)

```
## Helper Functions 


## Relevant Columns 

```{r}
relevant_cols <- (c(
    "test_id",
    "reference_number",
    "chemical_name",
    "Chemical_name",
    "test_cas",
    "species_number",
    "organism_habitat",
    "exposure_type",
    "control_type",
    "media_type",
    "genus",
    "common_name",
    "latin_name",
    "species",
    "ecotox_group",
    "obs_duration_mean",
    "obs_duration_unit",
    "endpoint",
    "effect",
    "measurement",
    "conc1_mean",
    "conc1_unit",
    "conversion_factors",
    "conc_ug_L", 
    "taxonomic.distance",
    "r", "R2", "intercept", "slope","MSE",
    "endpoint_type", "US-EPA New Chemical Categories", "CAS", "ASTER",
    "chemical_name",
    "ecotox_group",  
    "dtxsid", "epaName","formatted_cas"
  ))

```



Find tests with relevant cas and endpoints 
```{r}

# Join the three tables on relevant keys
ecotox_tests <- readRDS("~/Documents/repos/water100/data/ecotox_tests.rds")
sel_effect_codes <- c("MOR", "MORT", "GMOR", "PSUR", "SURV", 
                  "SVVS")


# View or store the result
candidate_tests <- ecotox_tests %>%
  filter(effect %in% sel_effect_codes)%>% 
  filter(media_type == "FW" | media_type == "FW/" ) #filter for freshwater only

  
#how many cas in candidate tests 
candidate_tests$test_cas %>% unique() %>% length()


```


```{r}
library(readxl)
moas <- read_excel(here("xlsx","envirotox_20230828155443.xlsx"), 
    sheet = "substance") %>% mutate(CAS = as.numeric(CAS))
  

```

```{r}


ecotox_tax_distance <- readRDS("~/Documents/repos/water100/data/ecotox_tax_distance.rds")

tax_dist = ecotox_tax_distance %>% select(c(common_name, latin_name, species_number, taxonomic.distance)) %>% distinct() 

summary_data <- candidate_tests %>% 
  left_join(tax_dist,by=join_by(species_number)) 

cas_salmonid_tests <- summary_data %>% filter(taxonomic.distance <= 4)


#join with ecotox data 
summary3 <- cas_salmonid_tests %>% 
  left_join(ecotox.chemicals %>% collect(),by=join_by(test_cas == cas_number)) %>% 
  #left_join(tax.distance,by=join_by(latin_name==latin_name.X)) %>% 
  mutate(test_cas = test_cas %>% as.numeric())

```

```{r}

#join summary data to moas 
summary4 <- summary3 %>% 
  left_join(moas,by=join_by(test_cas == CAS)) %>% 
  #add in endpoint descriptions 
  left_join(ecotox.endpoint_codes %>% collect() %>% 
            dplyr::select(c(code,endpoint_description = description))
            ,by=join_by(endpoint == code)) %>% 
  #remove trailing slash in some endpoint codes 
  mutate(endpoint = gsub("/|*", "",endpoint)) %>% 
  select(-chemical_name) %>% 
  #join with ecotox category 
  left_join(ecotox_cas_summary, by = join_by(test_cas == cas_number)) %>% select(any_of(relevant_cols)) 
```


Find tests and cas where lc50 for salmonids 
+ low vals for taxonomic.distance <5
+ high vals for taxonmic.distance <5
```{r}

endpoint_classifications <- read_excel(here("xlsx","endpoint_classifications.xlsx")) %>% 
  dplyr::select(c(endpoint,endpoint_type = Type,percent_val)) %>% filter(endpoint_type != "out" )

summary7 <- summary4 %>% left_join(endpoint_classifications,by=join_by(endpoint))%>% 
  dplyr::select(any_of(relevant_cols)) %>% filter(endpoint_type != "NA")


# 
# # Extracting the unique CAS values meeting the criteria
# cas_values <- filtered_data$test_cas

summary8 <- summary7 %>% 
  mutate(has_low = (endpoint_type == "low")) %>% 
  mutate(has_mid = (endpoint_type == "mid")) %>% 
  mutate(has_high = (endpoint_type == "high"))



filtered_cas <- summary8 %>%
  group_by(test_cas) %>%
  summarise(
    has_low_any = any(has_low == TRUE),
    has_mid_any = any(has_mid == TRUE),
    has_high_any = any(has_high == TRUE)
  )

filtered_cas <- filtered_cas %>%
  filter( 
    rowSums(filtered_cas %>% dplyr::select(c(has_low_any,has_mid_any,has_high_any))) == 3
  )

summary9 <- summary7 %>% 
  filter(test_cas %in% filtered_cas$test_cas) %>% 
  mutate(endpoint = gsub("/|[*]", "",endpoint)) %>% 
  left_join(ecotox_cas_summary.3 %>% select(
    c(cas_number,formatted_cas,epaName)
  ), by=join_by(test_cas == cas_number))

summary9 %>% clipr::write_clip()

library(readr)
library(here)
#write_rds(summary9,here("data",'cas_tests_taxdist.rds'))

test_filter <- summary9$test_id


test_outputs <- ecotox.tests %>% filter(test_id %in% test_filter) %>% collect()
write_rds(test_outputs,here('data','ecotox_w100_tests.rds'))

result_outputs <- ecotox.results %>% filter(test_id %in% test_filter) %>% collect()
write_rds(result_outputs %>% collect(),here('data','ecotox_w100_results.rds'))
#result_filter <- result_outputs$result_id

# dose_outputs <- ecotox.doses %>% filter(test_id %in% test_filter) %>% collect()
# 
# dose_link_outputs <- result_outputs %>% inner_join(ecotox.dose_response_links,by=join_by(result_id))
#   
#   result_outputsecotox.dose_response_links %>% filter(result_id %in% result_filter) %>% collect()
```
```{r}



```

